---
title: "Workday reporting"
execute:
  message: false
---

```{r}
#| label: get-file
#| message: false
library(tidyverse)
library(googledrive)
```

```{r}
#| label: token-cache-dir
#| message: false
# designate project-specific cache
cache_dir <- here::here(".secrets")
encrypted_token_path <- here::here(cache_dir, "drive-token.rds")

# set values in options
options(
  gargle_oauth_cache = cache_dir,
  gargle_oauth_email = TRUE,
  gargle_verbosity = "debug"
)
```

```{r}
#| eval: false
# Ben did this once -- no one else should need to do it!!!

# Step 1a: generate a secret key
secret_key <- gargle::secret_make_key()
# Step 1b: save the secret key as the GS4_KEY environment variable in `_environment.local`
# Sys.getenv("DRIVE_KEY")
# Sys.setenv(DRIVE_KEY = secret_key)

# Step 1c: Configure the secret key as a Secret on GitHub

# Step 2a: get a token and DO NOT CACHE IT
googledrive::drive_auth(email = "bbaumer@smith.edu", scopes = "drive.readonly", cache = FALSE)
# Step 2b: encrypt the token and write it to a file
gargle::secret_write_rds(
  x = drive_token(),
  path = encrypted_token_path,
  key = "DRIVE_KEY"
)
# Step 2c: commit the encrypted_token_path file to GitHub
```


```{r}
#| label: authorize-token
#| message: false
# now decrypt the token using the secret key. 
# This should work both locally AND on GitHub Actions!!
drive_auth(token = gargle::secret_read_rds(encrypted_token_path, "DRIVE_KEY"))
```


```{r}
read_excel_drive_cache <- function(.data, ...) {

  # Check if any files were found
  if (nrow(.data) == 0) {
    stop("No .xlsx files found in Google Drive.")
  }

  # Print the name of the file being downloaded
  message("Downloading the last alphabetical file:", .data$name, "\n")

  # 3. Download the selected file to a temporary location
  temp_file <- tempfile(fileext = ".xlsx")
  drive_download(
    file = .data, # A 'dribble' row can be passed directly to drive_download
    path = temp_file,
    overwrite = TRUE
  )

  # 4. Read the Excel file into R
  readxl::read_excel(temp_file, ...) |>
    janitor::clean_names()
}
```


```{r}
#| label: data-dir
#| message: false
# List only .xlsx files inside the "Chair" directory of the "SDS" Shared Drive
workday_data_dir <- drive_find(
  pattern = "Workday Data",
  type = "folder",
  shared_drive = "SDS"
)
```

```{r}
#| label: find-drive-files
#| message: false
workday_majors_sds <- workday_data_dir |>
  drive_ls() |>
  filter(str_detect(name, "workday_majors_sds")) |>
  drive_ls(
    pattern = "Department_majors",
    type = "xlsx"
  ) |>
  arrange(desc(name)) |>
  head(1) |>
  mutate(date_saved = parse_date(str_sub(name, 0, 8), "%Y%m%d"))

sds_majors <- workday_majors_sds |>
  read_excel_drive_cache() |>
  mutate(
    is_sds_major = str_detect(programs_of_study, "Statistical and Data Sciences Major"),
    is_mst_major = str_detect(programs_of_study, "Mathematical Statistics Major"),
    is_mth_major = str_detect(programs_of_study, "Mathematics Major"),
    is_sds_minor = str_detect(programs_of_study, "Statistical and Data Sciences Minor"),
    is_sts_minor = str_detect(programs_of_study, "Applied Statistics Minor"),
  )
```

## SDS Majors and minors

As of `r format(workday_majors_sds$date_saved, "%B %e, %Y")` were **`r nrow(sds_majors)`** SDS majors. 

```{r}
#| label: sds-majors
sds_majors |>
  group_by(class_year) |>
  summarize(
    sds_majors = sum(is_sds_major),
    mst_majors = sum(is_mst_major)
  ) |>
  arrange(class_year) |>
  janitor::clean_names(case = "title") |>
  knitr::kable()
```


## MST Majors

```{r}
#| message: false
workday_majors_mst <- workday_data_dir |>
  drive_ls() |>
  filter(str_detect(name, "workday_majors_mst")) |>
  drive_ls(
    pattern = "MST_Majors",
    type = "xlsx"
  ) |>
  arrange(desc(name)) |>
  head(1) |>
  mutate(date_saved = parse_date(str_sub(name, 0, 8), "%Y%m%d"))

mst_majors <- workday_majors_mst |>
  read_excel_drive_cache() |>
  mutate(
    is_sds_major = str_detect(major_s, "Statistical and Data Sciences Major"),
    is_mst_major = str_detect(major_s, "Mathematical Statistics Major"),
    is_mth_major = str_detect(major_s, "Mathematics Major")
  )
```

As of `r format(workday_majors_mst$date_saved, "%B %e, %Y")` were **`r nrow(mst_majors)`** MST majors. 

```{r}
#| label: mst-majors
mst_majors |>
  group_by(class_year) |>
  summarize(
    sds_majors = sum(is_sds_major),
    mst_majors = sum(is_mst_major)
  ) |>
  arrange(class_year) |>
  janitor::clean_names(case = "title") |>
  knitr::kable()
```



## Faculty advising loads

```{r}
#| message: false
workday_advisers <- workday_data_dir |>
  drive_ls() |>
  filter(str_detect(name, "workday_advising")) |>
  drive_ls(
    pattern = "Advisee_Count_Report",
    type = "xlsx",
  ) |>
  arrange(desc(name)) |>
  head(1)

y <- workday_advisers |>
  read_excel_drive_cache()
```

```{r}
sds_advisers <- y |>
  filter(majors_in_academic_unit > 0) |>
  select(-academic_units, -adviser_email, -adviser_id) |>
  mutate(
    total_wt_advisees = liberal_arts_advisees + majors_in_academic_unit + majors_out_of_academic_unit + 0.5 * (minor_advisees + concentration_advisees)
  )
sds_advisers |>
  arrange(desc(total_wt_advisees)) |>
  janitor::clean_names(case = "title") |>
  janitor::adorn_totals(where = "row") |>
  knitr::kable()
```

Among the `r (n <- nrow(filter(y, majors_in_academic_unit > 0)))` advising faculty members, the average faculty advising load is **`r round(nrow(sds_majors) / n, 2)`** SDS majors, **`r round(nrow(mst_majors) / n, 2)`** MST majors, and **`r round(sum(sds_advisers$total_wt_advisees) / n, 2)`** total weighted advisees.

## Enrollments

```{r}
#| message: false
workday_enrollments <- workday_data_dir |>
  drive_ls() |>
  filter(str_detect(name, "workday_enrollments")) |>
  drive_ls(
    pattern = "SDS Majors",
    type = "xlsx"
  ) |>
  arrange(desc(name)) |>
  head(1) |>
  mutate(date_saved = parse_date(str_sub(name, 0, 8), "%Y%m%d"))

enrollments <- workday_enrollments |>
  read_excel_drive_cache(skip = 1)

enrollments |>
  select(course_number, section_number, all_instructors, section_capacity, enrollment_count, waitlist_count) |>
  mutate(open_seats = section_capacity - enrollment_count - waitlist_count) |>
  arrange(course_number, section_number, open_seats) |>
  unique() |>
  janitor::clean_names(case = "title") |>
  knitr::kable()
```

## Capstone planning

```{r}
#| eval: false
sds_majors |>
  filter(is_sds_major, grepl("27|AC", class_year)) |>
  select(student, class_year) |>
  arrange(class_year, student) |>
  print(n = Inf)
```

